### 2024-10-25 scripts/05-exploratory_data_analysis.qmd
simulation revised
simulation revised

#### Preamble ####
# Purpose: Models... [...UPDATE THIS...]
# Author: Rohan Alexander [...UPDATE THIS...]
# Date: 11 February 2023 [...UPDATE THIS...]
# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
# License: MIT
# Pre-requisites: [...UPDATE THIS...]
# Any other information needed? [...UPDATE THIS...]

```{r}
#### Workspace setup ####
library(tidyverse)
library(rstanarm)

#### Read data ####
analysis_data <- read_csv(here::here("data/01-raw_data/raw_data.csv"))
```

```{r}
# Check for missing values in each column
missing_values <- sapply(analysis_data, function(x) sum(is.na(x)))
print(missing_values)
```

```{r}
# Get unique pollster IDs
unique_pollster_ids <- unique(analysis_data$pollster_id)
print(unique_pollster_ids)
```
```{r}
# Get unique question IDs
unique_question_ids <- unique(analysis_data$question_id)
print(unique_question_ids)
```
```{r}
unique_question_ids_str <- paste(unique(unique_question_ids), collapse = ",")
print(unique_question_ids_str)
```

```{r}
View(unique_question_ids_str)
```

```{r}
# Get the maximum and minimum values of numeric_grade
max_numeric_grade <- max(analysis_data$numeric_grade, na.rm = TRUE)
min_numeric_grade <- min(analysis_data$numeric_grade, na.rm = TRUE)
print(paste("Maximum numeric grade:", max_numeric_grade))
print(paste("Minimum numeric grade:", min_numeric_grade))
```

```{r}
# Plot histogram of numeric_grade
ggplot(analysis_data, aes(x = numeric_grade)) +
  geom_histogram(binwidth = 0.5, fill = "blue", color = "black") +
  theme_minimal() +
  labs(title = "Histogram of Numeric Grade", x = "Numeric Grade", y = "Count")
```
```{r}
# Plot histogram of pollscore
ggplot(analysis_data, aes(x = pollscore)) +
  geom_histogram(binwidth = 0.5, fill = "blue", color = "black") +
  theme_minimal() +
  labs(title = "Histogram of pollscore", x = "pollscore", y = "Count")
```
```{r}
# Get unique methodologys
unique_methodologys <- unique(analysis_data$methodology)
print(unique_methodologys)
```
```{r}
# Plot histogram of transparency_score
ggplot(analysis_data, aes(x = transparency_score)) +
  geom_histogram(binwidth = 0.5, fill = "blue", color = "black") +
  theme_minimal() +
  labs(title = "Histogram of transparency_score", x = "transparency_score", y = "Count")
```
```{r}
# Find the earliest start date and the latest end date
earliest_start_date <- min(analysis_data$start_date, na.rm = TRUE)
latest_end_date <- max(analysis_data$end_date, na.rm = TRUE)
print(paste("Earliest start date:", earliest_start_date))
print(paste("Latest end date:", latest_end_date))
```

```{r}
# Plot histogram of sample_size
ggplot(analysis_data, aes(x = sample_size)) +
  geom_histogram(binwidth = 0.5, fill = "blue", color = "black") +
  theme_minimal() +
  labs(title = "Histogram of sample_size", x = "sample_size", y = "Count")
```
```{r}
# Get unique population
unique_population <- unique(analysis_data$population)
print(unique_population)
```
```{r}
# Get unique race_id
unique_race_id <- unique(analysis_data$race_id)
print(unique_race_id)
```

```{r}
unique_race_id_str <- paste(unique(unique_race_id), collapse = ",")
clipr::write_clip(unique_race_id_str)
```

```{r}
# Get unique stage
unique_stage <- unique(analysis_data$stage)
print(unique_stage)
```

```{r}
# Get unique stage
unique_nationwide_batch <- unique(analysis_data$nationwide_batch)
print(unique_nationwide_batch)
```

```{r}
# Get unique ranked_choice_reallocated
unique_ranked_choice_reallocated <- unique(analysis_data$ranked_choice_reallocated)
print(unique_ranked_choice_reallocated)
```

```{r}
# Get unique hypothetical
unique_hypothetical <- unique(analysis_data$hypothetical)
print(unique_hypothetical)
```
```{r}
# Calculate the ratio of FALSE to TRUE in the hypothetical column
hypothetical_ratio <- table(analysis_data$hypothetical)
print(hypothetical_ratio)
false_to_true_ratio <- hypothetical_ratio["FALSE"] / hypothetical_ratio["TRUE"]
print(paste("False to True Ratio:", false_to_true_ratio))
```

```{r}
# Get unique party
unique_party <- unique(analysis_data$party)
print(unique_party)
```

```{r}
linear_model <- lm(pct ~ numeric_grade + pollscore + transparency_score + sample_size + partisan + methodology + population + race_id + stage + nationwide_batch + ranked_choice_reallocated + hypothetical, data = analysis_data)
summary(linear_model)
```