### 2024-11-01 scripts/05-exploratory_data_analysis.qmd
paper updated
paper updated

#### Preamble ####
# Purpose: Explore the analysis data and find representitive values
# Author: Haowei Fan, Fangning Zhang, Shaotong Li
# Date: 13 October 2024
# Contact: haowei.fan@mail.utoronto.ca
# License: MIT
# Pre-requisites: 02-analysis_data saved and loaded
# Any other information needed? None

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#### Workspace setup ####
library(tidyverse)
library(rstanarm)
library(patchwork)
raw_data <- read_csv(here::here("data/01-raw_data/raw_data.csv"))
```
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Get unique pollster IDs
unique_pollster_ids <- unique(raw_data$pollster_id)
print(unique_pollster_ids)
```
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Get unique question IDs
unique_question_ids <- unique(raw_data$question_id)
print(unique_question_ids)
```
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Get unique methodologys
unique_methodologys <- unique(raw_data$methodology)
print(unique_methodologys)
```
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Get unique population
unique_population <- unique(raw_data$population)
print(unique_population)
```
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Get unique race_id
unique_race_id <- unique(raw_data$race_id)
print(unique_race_id)
```
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Get unique ranked_choice_reallocated
unique_ranked_choice_reallocated <- unique(raw_data$ranked_choice_reallocated)
print(unique_ranked_choice_reallocated)
```
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Get unique hypothetical
unique_hypothetical <- unique(raw_data$hypothetical)
print(unique_hypothetical)
```
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Get unique party
unique_party <- unique(raw_data$party)
print(unique_party)
```
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Get the maximum and minimum values of numeric_grade
max_numeric_grade <- max(raw_data$numeric_grade, na.rm = TRUE)
min_numeric_grade <- min(raw_data$numeric_grade, na.rm = TRUE)
print(paste("Maximum numeric grade:", max_numeric_grade))
print(paste("Minimum numeric grade:", min_numeric_grade))
```
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Find the earliest start date and the latest end date
earliest_start_date <- min(raw_data$start_date, na.rm = TRUE)
latest_end_date <- max(raw_data$end_date, na.rm = TRUE)
print(paste("Earliest start date:", earliest_start_date))
print(paste("Latest end date:", latest_end_date))
```
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Calculate the ratio of FALSE to TRUE in the hypothetical column
hypothetical_ratio <- table(raw_data$hypothetical)
print(hypothetical_ratio)
false_to_true_ratio <- hypothetical_ratio["FALSE"] / hypothetical_ratio["TRUE"]
print(paste("False to True Ratio:", false_to_true_ratio))
```
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Find the range of pollscore in the original dataset
pollscore_range <- range(raw_data$pollscore, na.rm = TRUE)
print(paste("Pollscore range:", pollscore_range[1], "to", pollscore_range[2]))
```
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Plot histogram of numeric_grade
ggplot(raw_data, aes(x = numeric_grade)) +
  geom_histogram(binwidth = 0.5, fill = "blue", color = "black") +
  theme_minimal() +
  labs(title = "Histogram of Numeric Grade", x = "Numeric Grade", y = "Count")
```
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Plot histogram of pollscore
ggplot(raw_data, aes(x = pollscore)) +
  geom_histogram(binwidth = 0.5, fill = "blue", color = "black") +
  theme_minimal() +
  labs(title = "Histogram of pollscore", x = "pollscore", y = "Count")
```
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Plot histogram of transparency_score
ggplot(raw_data, aes(x = transparency_score)) +
  geom_histogram(binwidth = 0.5, fill = "blue", color = "black") +
  theme_minimal() +
  labs(title = "Histogram of transparency_score", x = "transparency_score", y = "Count")
```
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Plot histogram of sample_size
ggplot(raw_data, aes(x = sample_size)) +
  geom_histogram(binwidth = 0.5, fill = "blue", color = "black") +
  theme_minimal() +
  labs(title = "Histogram of sample_size", x = "sample_size", y = "Count")
```
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Find the top 3 candidates.
candidate_ranking <- raw_data %>%
  group_by(candidate_name) %>%
  summarize(
    poll_count = n(),                    
    avg_weighted_pct = mean(poll_count * pct, na.rm = TRUE)  
  ) %>%
  arrange(desc(avg_weighted_pct)) %>%    
  slice_head(n = 3)                       
print(candidate_ranking)
```